---
title: "STAT 685: Dr. Suojin Wang's Group"
subtitle: |
  | Modeling Seoul Bike Sharing Demand
author: "Nam Tran, Bai Zou"
date: "9/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setupAnchor, cache=TRUE, echo=FALSE, message=FALSE}
packages = c("tidyverse", "lubridate", "flipTime", "aTSA", "forecast", 
             "reshape2", "GGally", "MASS")
for(package in packages){
  if(!require(package,character.only=TRUE,quietly=TRUE)){
    install.packages(package,repos='http://cran.us.r-project.org')
    require(package,character.only=TRUE,quietly=TRUE)
  }
}
theme_update(plot.title = element_text(hjust = 0.5))

# Loading the Data
datPath = paste0(dirname(getwd()), "/data/SeoulBikeData.csv")
colNames = c("Date", "RentedBikeCount", "Hour", "Temp", "Humidity", 
             "WindSpeed", "Visibility", "DewPointTemp", "SolarRadiation",
             "Rainfall", "Snowfall", "Seasons", "Holiday", "FunctionalDay")
dat = read_csv(datPath, col_names = colNames, skip=1)

# Setting up Factors
dat$Seasons = as_factor(dat$Seasons)
dat$Holiday = as_factor(dat$Holiday)
dat$FunctionalDay = as_factor(dat$FunctionalDay)

# Creating DateTime that incorporates both Date and Hours and dropping Hour
dat$Date = AsDateTime(dat$Date) + hours(dat$Hour)
```

# Data Exploration  

## Qualtitative Variables

* The plots shows more rented bike count in non-holidays than holidays except for summer.

* If functional day is "no", there's no any bike rented.  

* Day of week is not making significant difference in rented bike count.

```{r qualtitativePlot, cache=TRUE, message=FALSE, fig.height=3, dependson='setupAnchor'}
# adding day of time attributes
dat$DayOfWeek <- weekdays(dat$Date)

# plotting by Seasons and Holiday
dat %>% 
  ggplot(aes(x=Seasons, y=RentedBikeCount, fill=Holiday)) + 
  geom_boxplot() +
  scale_fill_brewer(palette="Paired") +
  labs(y="Rented Bike Count", x="Seasons", title="Rented Bike Count by Seasons and Holiday") 

# plotting by Seasons and FunctionalDay
dat %>% 
  ggplot(aes(x=Seasons, y=RentedBikeCount, fill=FunctionalDay)) + 
  geom_boxplot() +
  scale_fill_brewer(palette="Paired") +
  labs(y="Rented Bike Count", x="Seasons", title="Rented Bike Count by Seasons and FunctionalDay")

# plotting by Seasons and DayOfWeek
dat %>% 
  ggplot(aes(x=DayOfWeek, y=RentedBikeCount, fill=Seasons)) + 
  geom_boxplot() +
  scale_fill_brewer(palette="Paired") +
  labs(y="Rented Bike Count", x="DayOfWeek", title="Rented Bike Count by Seasons and DayOfWeek")
```


## Quantitative Variables

* The covariance matrix shows Temp, Hour has relatively higher correlation with RentedBikeCount (>0.4).

* DewPointTemp and SolarRadiation have correlation greater than 0.2. 

* Temp and DewPointTemp are highly correlated (0.9). 

* **No clear linear relationship can be identified between response variable and quantitative Variables**

```{r quantitativePlot, cache=TRUE, message=FALSE, fig.height=3, dependson='setupAnchor'}
quantitative_var = c("Hour", "Temp", "Humidity", "WindSpeed", "Visibility", "DewPointTemp", 
             "SolarRadiation", "Rainfall", "Snowfall")

# check covariance  
cor_matrix = cor(dat[c("RentedBikeCount", quantitative_var)])
cor_matrix2 = melt(cor_matrix)

cor_matrix2 %>% 
  ggplot(aes(x=Var1, y=Var2, fill=abs(value))) + 
    geom_tile() +
    theme(axis.text.x = element_text(angle = 90))

cor_matrix = cor_matrix[order(abs(cor_matrix[,1]), decreasing=TRUE),]
data.frame(cor_matrix[2:10, 1])
```
```{r quantitativePlot2, cache=TRUE, message=FALSE, fig.height=6, dependson='setupAnchor'}
# scatter plot matrix
select_var = rownames(cor_matrix)[1:5]
p = ggpairs(dat[select_var], 
            aes(colour = dat$Seasons, alpha = 0.2))
show(p)
```

# Simple Linear Regression Fit

* Hour is highly related to RentedBikeCount, but not linearly related. Using Hour as qualtitative variable improves 10% in R-squared. 

* Adding second order for Temp doesn't bring significant improvement. 

* Even with some modification on variables, the simple linear fit is still not very good with R-squared below 70%. The residual plots are showing 'v' shaped pattern that needs further investigation. 


## Baseline

```{r simpleFit, cache=TRUE, message=FALSE, fig.height=5, dependson='setupAnchor'}
fit_dat1 = dat[, -1]
# Fit the full model 
full.model <- lm(RentedBikeCount ~., data = fit_dat1)
# Stepwise regression model
step.model <- stepAIC(full.model, direction = "both", 
                      trace = FALSE)
summary(step.model)

# Residual plots
par(mfrow = c(2, 2)) 
plot(step.model) 
```


## Treat Hour as Qualtitative Variables

```{r fit2, cache=TRUE, message=FALSE, fig.height=5, dependson='setupAnchor'}
fit_dat2 = dat[, -1]
fit_dat2$Hour = as.factor(fit_dat2$Hour)
# Fit the full model 
full.model2 <- lm(RentedBikeCount ~., data = fit_dat2)
# Stepwise regression model
step.model2 <- stepAIC(full.model2, direction = "both", 
                      trace = FALSE)
summary(step.model2)

# Residual Plots
par(mfrow = c(2, 2))  # Split the plotting panel into a 2 x 2 grid
plot(step.model2)  # Plot the model information
```

## Adding 2nd Order of Temp

```{r fit3, cache=TRUE, message=FALSE, fig.height=5, dependson='setupAnchor'}
fit_dat3 = dat[, -1]
fit_dat3$Hour = as.factor(fit_dat3$Hour)

# mean temp
fit_dat3$Temp2 = fit_dat3$Temp ^ 2

# Fit the full model 
full.model3 <- lm(RentedBikeCount ~., data = fit_dat3)
# Stepwise regression model
step.model3 <- stepAIC(full.model3, direction = "both", 
                      trace = FALSE)
summary(step.model3)

# Residual Plots
par(mfrow = c(2, 2))  # Split the plotting panel into a 2 x 2 grid
plot(step.model3)  # Plot the model information
```

